Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-02-03 paper/elsarticle/main.tex
grammarly

			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any further tuning or interventions of the control engineer during operation. Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. It provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behaviour. To ensure offset-free reference tracking control, the built-in integrator was included in the explicit MPC design. The asymmetric behaviour of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning process. The designed self-tunable controller improved control performance by decreasing integral square error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (not tuned) controller. 
		%\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%%Research highlights
		%\begin{highlights}
			%% 3-5 highlights, every one should not exceed 85 characters including spaces
			%\item Self-tunable control technique does not need any intervention during control
			%\item Tuning based on the operating conditions compensates the nonlinear plant behaviour
			%\item Implementation of the proposed tuning method on a laboratory heat exchanger 
			%\item Control performance improved compared to utilizing only one explicit MPC
			%\end{highlights}
		%
		%\begin{keyword}
			%tunable explicit MPC \sep self-tunable technique \sep tuning parameter \sep heat exchanger \sep control performance
			%%% keywords here, in the form: keyword \sep keyword
			%
			%%% PACS codes here, in the form: \PACS code \sep code
			%
			%%% MSC codes here, in the form: \MSC code \sep code
			%%% or \MSC[2008] code \sep code (2000 is the default)
			%
			%\end{keyword}
		%
		%\end{frontmatter}

@@

%	TODO: Vyhnut sa pouzivaniu "we", "our",...
%	
%	TODO: dopisat zaciatok uvodu o vymennikoch tepla, ladeni regulatorov vseobecne, gain-scheduling atd (neskor je pokryte MPC, MPC riadenie vymennikov tepla, eMPC, tunable eMPC a self-tunable)

@@

	The heat exchangers in their numerous variants integrated into the many industrial plants represent energy-demanding plants. From the control viewpoint, the controller design taking into account the nonlinear and asymmetric behaviour of the heat exchangers is a challenging task, see~\cite{RL20}. Although the conventional and widely-used PID controllers are robust and easy to tune, their control performance may not be sufficient when comes to energy savings. Various extensions built above the well-tuned PID controller were developed to compensate for the non-linear and asymmetric behaviour, often affected by the additional impact of the uncertain parameters originated e.g., in the fouling~\cite{MT19}. Such widely-used control strategies include, e.g., the robust control~\cite{DO15,WY18}, the gain-scheduling, multi-zone control, see~\cite{MF08,Liptak}, and references therein.

	One of the promising control strategies addressing all these issues in an optimal way came with the formation and tuning of the model predictive control (MPC), e.g., see~\cite{Morari_MPC}. MPC provides optimal control input based on the minimization of a specified cost function while considering a model of the system. Compared to linear quadratic controllers (LQR)~\cite{LQR}, model predictive control also includes constraints on the control input or process variables~\cite{Maciejowski_MPC}, and additional saturation is not necessary. Moreover, as the optimization problem is solved in every control step, MPC represents a receding horizon control policy~\cite{receding_horizon} having a significant benefit mainly in the terms of disturbance rejection. The model predictive control was intensively investigated in connection with heat exchangers. In~\cite{Vinaya_HE_MPC}, the authors developed a model predictive control for a shell and tube heat exchanger. In~\cite{Oravec_energy}, the control performance of the robust MPC implemented for a plate heat exchanger with soft constraints was analyzed. Four robust control strategies were presented and compared in~\cite{Oravec_HE_ATE}. A two-level control structure was applied on a heat exchanger network in~\cite{Gonzales_HE_MPC}. In this study, the low level of control was ensured by MPC and the high level by a supervisory online optimiser.

@@

	The paper~\cite{self_tunable} pushes the idea of tunable explicit MPC further and deals with the issues of practical industrial-oriented implementation. The procedure of the self-tunable controller technique is presented. The controller's aggressivity is tuned based on the difference between the reference value and the steady state corresponding to the model linearization point. Tuning based on the distance from the steady-state operating point represented a way how to compensate for the system's nonlinear behaviour.

	This work directly extends our results presented in~\cite{self_tunable}, where the basic principles of the self-tunable approximated explicit MPC were introduced. In this paper, a novel method of self-tuning parameter setup is introduced. Compared to~\cite{self_tunable}, the self-tuning method is based on the size of the reference step change. Moreover, the idea of further scaling the tuning parameter is elaborated. The interval of the values of the self-tuning parameter is split at some certain value, where each part of the interval corresponds to the specific operating conditions defined by the control engineer. In such a way, e.g., the system's asymmetric behaviour is compensated. Finally, to investigate the benefits of the proposed approach, the proposed self-tuning control policy was implemented to control a laboratory-scaled counter-current plate heat exchanger. This work provides the control performance evaluation and analysis using one non-tunable controller.

	The rest of the paper is organized as follows. First, the theoretical backgrounds are presented in Section~\ref{sec:preliminaries}, where the explicit MPC, the approximated tunable explicit MPC, and existing self-tunable methods are briefly elaborated. Then, the novel proposed method of self-tunable procedure is explained in detail in Section~\ref{sec:methodology}. Finally, the experimental results of the self-tuning controller implementation on a heat exchanger are discussed in Section~\ref{sec:results}, followed by the main conclusions in Section~\ref{sec:conclusion}.

@@

	Explicit model predictive control~\cite{Bemporad_automatica} utilizes a parametric solution of the model predictive control introducing its application range towards the systems with fast dynamics. Moreover, the explicit solution enables providing rigorous analysis and certification of the closed-loop system stability, constraints satisfaction, etc. As the explicit solution is available, real-time solving of the optimization problem in every control step is omitted. As this work deals with industrial-oriented implementation, consider the optimization problem in the following form:
	\begin{subequations}
		\label{eq:mpc_problem}
		\begin{eqnarray}
			\label{eq:mpc_problem_cost}
			\min_{u_0,u_{1},\ldots,u_{N-1}} &~& \! \sum_{k=0}^{N-1} \! \left( (y_k-y_\mathrm{ref})^{\intercal} Q_\mathrm{y} (y_k-y_\mathrm{ref}) + u_{k}^{\intercal} R u_{k} + x_{\mathrm{I},k}^{\intercal} Q_\mathrm{x} x_{\mathrm{I},k} \right)  \\
			\label{eq:mpc_problem_prediction_model_x}
			\mathrm{s.t.\!:} &~& \widetilde{x}_{k+1} = \widetilde{A}\,\widetilde{x}_{k} + \widetilde{B}\,u_{k}, \\
			\label{eq:mpc_problem_prediction_model_y}
			&~& y_{k} = \widetilde{C}\,\widetilde{x}_{k}, \\
			\label{eq:mpc_problem_input_constraints}
			&~& u_{k} \in \mathcal{U}, \\
			\label{eq:mpc_problem_state_constraints}
			&~& y_{k} \in \mathcal{Y}, \\
			\label{eq:mpc_problem_initial_coindition}
			&~& \widetilde{x}_{0} = \theta, \\
			\label{eq:mpc_problem_k_range}
			&~& k = 0,1,\ldots, N-1,
		\end{eqnarray}
	\end{subequations}
	where $k$ denotes the step of the prediction horizon $N$. 
	%
	To obtain the offset-free control results, the built-in integrator was included in the state-space model, e.g., see~\cite{Ruscio_MPC_integral}. 
	The prediction model in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} has the form of augmented linear time-invariant (LTI) system for a given augmented state matrix $\widetilde{A} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\widetilde{\mathrm{x}}}}$, augmented input matrix $\widetilde{B} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\mathrm{u}}}$ and augmented output matrix $\widetilde{C} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\widetilde{\mathrm{x}}}}$. 
	%
	Variables $\widetilde{x} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}}}$, $u \in \mathbb{R}^{n_{\mathrm{u}}}$, $y \in \mathbb{R}^{n_{\mathrm{y}}}$ are vectors of corresponding augmented system states, control inputs and system outputs, respectively. 
	%
	The sets $\mathcal{U} \subseteq \mathbb{R}^{n_{\mathrm{u}}}$, $\mathcal{Y} \subseteq \mathbb{R}^{n_{\mathrm{y}}}$ are convex polytopic sets of physical constraints on inputs and outputs, respectively. The penalty matrix $Q_\mathrm{y} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$ penalizes the squared control error, i.e., the deviation between the output and output reference value $y_\mathrm{ref}$. The matrix $R \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{u}}}$ penalizes the squarred value of control inputs. 
	%
	The value of integrator is also penalized in the cost function with the penalty matrix $Q_\mathrm{x} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$. 
	%
	The parameter $\theta \in \Theta$ in Eq.~\eqref{eq:mpc_problem_initial_coindition} represents the initial condition of the optimization problem for which it is parametrically pre-computed.

	The augmented model of the controlled system with the built-in integrator in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} us rewritten as follows:
	\begin{subequations}
		\begin{eqnarray} 
			\label{eq:mpc_augmented_model_x} 
			\widetilde{x}_{k+1} &=& \begin{bmatrix} x_{k+1} \\ x_{\mathrm{I},k+1}\end{bmatrix} = \begin{bmatrix} A & \textit{0} \\ -T_\mathrm{s} C & I \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix} + \begin{bmatrix} B \\ I \end{bmatrix} u_{k}, \\
			\label{eq:mpc_augmented_model_y}
			y_k &=& \begin{bmatrix} C & \textit{0} \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix},
		\end{eqnarray}
	\end{subequations}
	where $x_{\mathrm{I}} \in \mathbb{R}^{n_{\mathrm{y}}}$ is the integral of the control error, $T_\mathrm{s}$ denotes the sampling time and matrices $A$, $B$, $C$ are well-known the state-space matrices that form the augmented LTI model. As the consequence of this extension and penalization in the cost function in Eq.~\eqref{eq:mpc_problem_cost}, not only the control error is penalized, but also an integrated value, which leads to analogous offset-free reference tracking results as incorporating an integral part in PID controller.

@@

	Note, many other formulations of the optimization problems for the explicit MPC design, were mainly in the terms of the definition of the cost functions in Eq.~\eqref{eq:mpc_problem_cost}. Also, the incremental (velocity) formulation of the state-space model is common, but leads to further extension of the vector of parameters $\theta$ and therefore also the complexity of the explicit MPC controller increases. Another option for offset-free tracking is introducing disturbance modelling and estimation. For such an overview see e.g.~\cite{Klauco_mpc}

@@

	The aggressivity of the controller and the whole nature of the control is influenced by appropriate fine-tuning of the penalty matrices in the optimization problem in Eq.~\eqref{eq:mpc_problem}. When the multi-parametric QP (mp-QP) problem is precomputed offline to obtain the corresponding parametric solution, it is not possible to tune the controller afterwards without trading off a significant increase in the controller complexity or the performance loss. As the operation conditions and requirements of controller setup may differ throughout the control, the ability to adjust the controller's aggressivity can be very beneficial.

@@

	Let us consider the penalty matrices in cost function in Eq.~\eqref{eq:mpc_problem_cost}. The penalty matrices are scaled in the following way:
	\begin{subequations}
		\label{eq:tunable_matrices}
		\begin{eqnarray}
			\label{eq:tunable_R}
			&~& R(k) = (1-\rho(k)) \, R_\mathrm{L} + \rho(k) \, R_\mathrm{U}, \\
			\label{eq:tunable_Qx}
			&~& Q_\mathrm{x}(k) = (1-\rho(k)) \, Q_\mathrm{x,L} + \rho(k) \, Q_\mathrm{x,U}, \\
			\label{eq:tunable_Qy}
			&~& Q_\mathrm{y}(k) = (1-\rho(k)) \, Q_\mathrm{y,L} + \rho(k) \, Q_\mathrm{y,U},
		\end{eqnarray}
	\end{subequations}
	where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in Eq.~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in Eq.~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance subject to the specific system and systematically tune all of the penalty matrices.

@@

	The concept of explicit MPC tuning is applicable to a wide class of MPC design formulations, based on the current specific needs. Without loss of generality, hereafter, let us consider the penalty matrices of cost function in Eq.~\eqref{eq:mpc_problem_cost}, as it is necessary to satisfy offset-free reference tracking.

@@

	The need for real-time controller tuning often arises from tracking a time-varying piece-wise constant (PWC) reference. The work~\cite{self_tunable} focuses on adjusting the penalty matrix when the reference value is changed. The further the reference value is from the steady state, the more aggressively the controller is tuned. The idea behind the suggested scaling lies in compensation for the nonlinear behaviour of the system.

	Consider a single-input and single-output (SISO) system or a multiple-inputs and multiple-outputs (MIMO) system with completely decoupled pairs of the control inputs and the system outputs. 
	Then, the procedure of tuning the controller is based on evaluating the different operating points between the current value of the reference and the system steady-state value. This deviation is considered to scale the value of control action. First, the maximal admissible absolute value of the reference is defined. Analogous to the reference trajectory preview concept of MPC design, this value can be determined based on the general knowledge of the expected future reference values. Another suggestion is to fit the maximal deviation $d_{\max}$ based on the constraints on system outputs: 
	\begin{eqnarray}
		\label{eq:d_max}
		d_{\max} = \max(\vert y_{\min} \vert, y_{\max}),
	\end{eqnarray}
	where $y_{\min}$ and $y_{\max}$ are respectively lower and upper bound on the output variable in the deviation form, ie., zero corresponds to the system steady-state value. Using the information about the maximal possible deviation $d_{\max}$, the tuning parameter $\rho$ can be calculated as the ratio between the current reference value and the maximal deviation:  
	\begin{eqnarray}
		\label{eq:rho_self_tunable}
		\rho = \frac{\vert y_{\mathrm{ref}(k)} \vert}{d_{\max}}.
	\end{eqnarray}

@@

	Note, the relations in Eq.~\eqref{eq:rho_self_tunable} and Eq.~\eqref{eq:rho_max} operate with the absolute value of the reference. It is not taken into account, whether the reference change was positive or negative with respect to the system steady-state value placed in the origin, i.e., whether it changed upwards or downwards. As many plants have asymmetric behaviour, the positivity or negativity of the reference change could be considered in the controller tuning to improve the control performance.

@@

	\subsection{Tuning parameter based on the size of reference change}
	\label{sec:self_tunable_delta_ref}
	The approach of self-tunable explicit MPC in~\cite{self_tunable} suggested tuning based on the current reference value distance from the steady state. The aim is to compensate for the nonlinear behaviour of a system when using a simple linear prediction model. This work provides also another useful way of the real-time evaluation of the tuning parameter $\rho$ based on the size of reference change. When different sizes of reference step changes are made and the behaviour of the closed-loop system is varying, it can be beneficial to include the size of the reference step change in the tuning procedure.

@@

	\subsection{Self-tunable technique for systems with asymmetric behaviour}
	\label{sec:self_tunable_rho_scaling}

	This paper provides a further extension of the self-tuning method proposed in~\cite{self_tunable}. The suggested technique of tuning is suitable, e.g., for systems with asymmetric behaviour, but can be used in any application, where ``simple'' tuning in the whole range of tuning parameter $\rho$ is not sufficient.

@@

	%% TODO: DEF RULE/CONDITION GAMMA
	\begin{definition}[Decision function]
		\label{def:gamma}
		For given interval of tuning parameter $\rho$, $0 \leq \rho \leq 1$, let $\rho_{\mathrm{s}}$, $0 < \rho_{\mathrm{s}} < 1$ be a boundary value splitting the interval into two parts. Let $\gamma: \mathbb{R} \rightarrow \mathbb{R}$ be an arbitrary function such that $0 \leq \gamma \leq 1$ holds. Then the decision function $\gamma$ is constructed to assign its value either $\gamma \leq \rho_{\mathrm{s}}$ or $\gamma > \rho_{\mathrm{s}}$.
	\end{definition}
	\textcolor{red}{TODO: replace in the line above: $\gamma \ge \rho_{\mathrm{s}}$?}\\
	Various decision functions $\gamma$ can be considered. In this work, the decision functions according to Eq.~\eqref{eq:rho_max} and Eq.~\eqref{eq:rho_delta_ref} are considered, while Eq.~\eqref{eq:rho_delta_ref} was implemented in the experimental case study.

@@

	For better insight of the proposed technique, the procedure of self-tuning is depicted in Fig.~\ref{fig:tuning}.

>>>>

		The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any further tuning or interventions of the control engineer during operation. Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. It provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. To ensure offset-free reference tracking control, the built-in integrator was included in the explicit MPC design. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning process. The designed self-tunable controller improved control performance by decreasing integral squared error, maximal overshoots/undershoots and settling time compared to the conventional control strategy based on a single (not tuned) controller. 
		%\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%%Research highlights
		%\begin{highlights}
			%% 3-5 highlights, every one should not exceed 85 characters including spaces
			%\item Self-tunable control technique does not need any intervention during control
			%\item Tuning based on the operating conditions compensates the nonlinear plant behaviour
			%\item Implementation of the proposed tuning method on a laboratory heat exchanger 
			%\item Control performance improved compared to utilizing only one explicit MPC
			%\end{highlights}
		%
		%\begin{keyword}
			%tunable explicit MPC \sep self-tunable technique \sep tuning parameter \sep heat exchanger \sep control performance
			%%% keywords here, in the form: keyword \sep keyword
			%
			%%% PACS codes here, in the form: \PACS code \sep code
			%
			%%% MSC codes here, in the form: \MSC code \sep code
			%%% or \MSC[2008] code \sep code (2000 is the default)
			%
			%\end{keyword}
		%
		%\end{frontmatter}

@@

@@

	The heat exchangers, in their numerous variants integrated into many industrial plants, represent energy-demanding plants. From the control viewpoint, the controller design taking into account the nonlinear and asymmetric behavior of the heat exchangers is a challenging task, see~\cite{RL20}. Although the conventional and widely-used proportional–integral–derivative (PID) controllers are robust and easy to tune, their control performance may not be sufficient when it comes to energy savings. Various extensions built above the well-tuned PID controller were developed to compensate for the nonlinear and asymmetric behavior, often affected by the additional impact of the uncertain parameters originated, e.g., in the fouling~\cite{MT19}. Such widely-used control strategies include, e.g., the robust control~\cite{DO15,WY18}, the gain-scheduling, multi-zone control, see~\cite{MF08}, and references therein.

	TODO: nepouzivat nasobne referencie ako [4,5] vyssie

	One of the promising control strategies addressing all these issues in an optimal way came with the formation and tuning of the model predictive control (MPC), e.g., see~\cite{Morari_MPC}. MPC provides optimal control input based on the minimization of a specified cost function while considering a model of the system. Compared to linear quadratic controllers (LQR)~\cite{LQR}, model predictive control also includes constraints on the control input or process variables~\cite{Maciejowski_MPC}, and additional saturation is not necessary. Moreover, as the optimization problem is solved in every control step, MPC represents a receding horizon control policy~\cite{receding_horizon}, having a significant benefit mainly in the terms of disturbance rejection. The model predictive control was intensively investigated in connection with heat exchangers. In~\cite{Vinaya_HE_MPC}, the authors developed a model predictive control for a shell and tube heat exchanger. In~\cite{Oravec_energy}, the control performance of the robust MPC implemented for a plate heat exchanger with soft constraints was analyzed. Four robust control strategies were presented and compared in~\cite{Oravec_HE_ATE}. A two-level control structure was applied on a heat exchanger network in~\cite{Gonzales_HE_MPC}, where the low level of control was ensured by MPC and the high level by a supervisory online optimizer.

@@

	The paper~\cite{self_tunable} pushes the idea of tunable explicit MPC further and deals with the issues of practical industrial-oriented implementation. The procedure of the self-tunable controller technique is presented. The controller's aggressivity is tuned based on the difference between the reference value and the steady state corresponding to the model linearization point. Tuning based on the distance from the steady-state operating point represented a way how to compensate for the system's nonlinear behavior.

	This work directly extends our results presented in~\cite{self_tunable}, where the basic principles of the self-tunable approximated explicit MPC were introduced. In this paper, a novel method of self-tuning parameter setup is introduced. Compared to~\cite{self_tunable}, the self-tuning method is based on the size of the reference step change. Moreover, the idea of further scaling the tuning parameter is elaborated. The interval of the values of the self-tuning parameter is split at some certain value and each part of the interval corresponds to the specific operating conditions defined by the control engineer. In such a way, e.g., the system's asymmetric behavior is compensated. Finally, to investigate the benefits of the proposed approach, the proposed self-tuning control policy was implemented to control a laboratory-scaled counter-current plate heat exchanger. This work provides the control performance evaluation and analysis using the self-tunable controller compared to the boundary explicit MPCs.

	The paper is organized as follows. First, the theoretical backgrounds are presented in Section~\ref{sec:preliminaries}, where the explicit MPC, the approximated tunable explicit MPC, and existing self-tunable methods are briefly elaborated. Then, the novel proposed method of self-tunable procedure is explained in detail in Section~\ref{sec:methodology}. Finally, the experimental results of the self-tuning controller implementation on a heat exchanger are discussed in Section~\ref{sec:results}, followed by the main conclusions in Section~\ref{sec:conclusion}.

@@

	Explicit model predictive control~\cite{Bemporad_automatica} utilizes a parametric solution of the model predictive control introducing its application range towards the systems with fast dynamics. Moreover, the explicit solution enables providing rigorous analysis and certification of the closed-loop system stability, constraints satisfaction, etc. As the explicit solution is available, real-time solving of the optimization problem in every control step is omitted. As this work deals with industrial-oriented implementation, let us consider the optimization problem in the following form:
	\begin{subequations}
		\label{eq:mpc_problem}
		\begin{eqnarray}
			\label{eq:mpc_problem_cost}
			\min_{u_0,u_{1},\ldots,u_{N-1}} &~& \! \sum_{k=0}^{N-1} \! \left( (y_k-y_\mathrm{ref})^{\intercal} Q_\mathrm{y} (y_k-y_\mathrm{ref}) + u_{k}^{\intercal} R u_{k} + x_{\mathrm{I},k}^{\intercal} Q_\mathrm{x} x_{\mathrm{I},k} \right)  \\
			\label{eq:mpc_problem_prediction_model_x}
			\mathrm{s.t.\!:} &~& \widetilde{x}_{k+1} = \widetilde{A}\,\widetilde{x}_{k} + \widetilde{B}\,u_{k}, \\
			\label{eq:mpc_problem_prediction_model_y}
			&~& y_{k} = \widetilde{C}\,\widetilde{x}_{k}, \\
			\label{eq:mpc_problem_input_constraints}
			&~& u_{k} \in \mathcal{U}, \\
			\label{eq:mpc_problem_state_constraints}
			&~& y_{k} \in \mathcal{Y}, \\
			\label{eq:mpc_problem_initial_coindition}
			&~& \widetilde{x}_{0} = \theta, \\
			\label{eq:mpc_problem_k_range}
			&~& k = 0,1,\ldots, N-1,
		\end{eqnarray}
	\end{subequations}
	where $k$ denotes the step of the prediction horizon $N$. 
	%
	To obtain the offset-free control results, the built-in integrator was included in the state-space model, e.g., see~\cite{Ruscio_MPC_integral}. 
	The prediction model in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} has the form of augmented linear time-invariant (LTI) system for a given augmented state matrix $\widetilde{A} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\widetilde{\mathrm{x}}}}$, augmented input matrix $\widetilde{B} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\mathrm{u}}}$ and augmented output matrix $\widetilde{C} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\widetilde{\mathrm{x}}}}$. 
	%
	Variables $\widetilde{x} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}}}$, $u \in \mathbb{R}^{n_{\mathrm{u}}}$, $y \in \mathbb{R}^{n_{\mathrm{y}}}$ are vectors of corresponding augmented system states, control inputs, and system outputs, respectively. 
	%
	The sets $\mathcal{U} \subseteq \mathbb{R}^{n_{\mathrm{u}}}$, $\mathcal{Y} \subseteq \mathbb{R}^{n_{\mathrm{y}}}$ are convex polytopic sets of physical constraints on inputs and outputs, respectively. The penalty matrix $Q_\mathrm{y} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$ penalizes the squared control error, i.e., the deviation between the output and output reference value $y_\mathrm{ref}$. The matrix $R \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{u}}}$ penalizes the squarred value of control inputs. 
	%
	The value of integrator is also penalized in the cost function with the penalty matrix $Q_\mathrm{x} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$. 
	%
	The parameter $\theta \in \Theta$ in Eq.~\eqref{eq:mpc_problem_initial_coindition} represents the initial condition of the optimization problem for which it is parametrically pre-computed.

	The augmented model of the controlled system with the built-in integrator in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} us rewritten as follows:
	\begin{subequations}
		\begin{eqnarray} 
			\label{eq:mpc_augmented_model_x} 
			\widetilde{x}_{k+1} &=& \begin{bmatrix} x_{k+1} \\ x_{\mathrm{I},k+1}\end{bmatrix} = \begin{bmatrix} A & \textit{0} \\ -T_\mathrm{s} C & I \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix} + \begin{bmatrix} B \\ I \end{bmatrix} u_{k}, \\
			\label{eq:mpc_augmented_model_y}
			y_k &=& \begin{bmatrix} C & \textit{0} \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix},
		\end{eqnarray}
	\end{subequations}
	where $x_{\mathrm{I}} \in \mathbb{R}^{n_{\mathrm{y}}}$ is the integral of the control error, $T_\mathrm{s}$ denotes the sampling time, and matrices $A$, $B$, $C$ are well-known the state-space matrices that form the augmented LTI model. As the consequence of this extension and penalization in the cost function in Eq.~\eqref{eq:mpc_problem_cost}, not only the control error is penalized, but also an integrated value, which leads to analogous offset-free reference tracking results as incorporating an integral part in PID controller.

@@

	Note, many other formulations of the optimization problems for the explicit MPC design, were mainly in the terms of the definition of the cost functions in Eq.~\eqref{eq:mpc_problem_cost}. Also, the incremental (velocity) formulation of the state-space model is common, but leads to further extension of the vector of parameters $\theta$, and therefore also the complexity of the explicit MPC controller increases. Another option for offset-free tracking is introducing disturbance modeling and estimation. For such an overview see e.g.~\cite{Klauco_mpc}

@@

	The aggressivity of the controller and the whole nature of the control is influenced by appropriate fine-tuning of the penalty matrices in the optimization problem in Eq.~\eqref{eq:mpc_problem}. When the multi-parametric QP (mp-QP) problem is precomputed offline to obtain the corresponding parametric solution, it is not possible to tune the controller afterward without trading off a significant increase in the controller complexity or the performance loss. As the operating conditions and requirements of controller setup may differ throughout the control, the ability to adjust the controller's aggressivity can be very beneficial.

@@

	Let us consider the penalty matrices in the cost function in Eq.~\eqref{eq:mpc_problem_cost}. The penalty matrices are scaled in the following way:
	\begin{subequations}
		\label{eq:tunable_matrices}
		\begin{eqnarray}
			\label{eq:tunable_R}
			&~& R(k) = (1-\rho(k)) \, R_\mathrm{L} + \rho(k) \, R_\mathrm{U}, \\
			\label{eq:tunable_Qx}
			&~& Q_\mathrm{x}(k) = (1-\rho(k)) \, Q_\mathrm{x,L} + \rho(k) \, Q_\mathrm{x,U}, \\
			\label{eq:tunable_Qy}
			&~& Q_\mathrm{y}(k) = (1-\rho(k)) \, Q_\mathrm{y,L} + \rho(k) \, Q_\mathrm{y,U},
		\end{eqnarray}
	\end{subequations}
	where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in Eq.~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in Eq.~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance subject to the specific system and systematically tune all of the penalty matrices.

@@

	The concept of explicit MPC tuning is applicable to a wide class of MPC design formulations, based on the current specific needs. Without loss of generality, hereafter, let us consider the penalty matrices of the cost function in Eq.~\eqref{eq:mpc_problem_cost}, as it is necessary to satisfy offset-free reference tracking.

@@

	The need for real-time controller tuning often arises from tracking a time-varying piece-wise constant (PWC) reference. The work~\cite{self_tunable} focuses on adjusting the penalty matrix when the reference value is changed. The further the reference value is from the steady state, the more aggressively the controller is tuned. The idea behind the suggested scaling lies in compensation for the nonlinear behavior of the system.

	Consider a single-input and single-output (SISO) system or a multiple-inputs and multiple-outputs (MIMO) system with completely decoupled pairs of the control inputs and the system outputs. 
	Then, the procedure of tuning the controller is based on evaluating the different operating points between the current value of the reference and the system steady-state value. This deviation is considered to scale the value of control action. First, the maximal admissible absolute value of the reference is defined. Analogous to the reference trajectory preview concept of MPC design, this value can be determined based on the general knowledge of the expected future reference values. Another suggestion is to fit the maximal deviation $d_{\max}$ based on the constraints on system outputs: 
	\begin{eqnarray}
		\label{eq:d_max}
		d_{\max} = \max(\vert y_{\min} \vert, y_{\max}),
	\end{eqnarray}
	where $y_{\min}$ and $y_{\max}$ are respectively lower and upper bound on the output variable in the deviation form, i.e., zero corresponds to the system steady-state value. Using the information about the maximal possible deviation $d_{\max}$, the tuning parameter $\rho$ can be calculated as the ratio between the current reference value and the maximal deviation:  
	\begin{eqnarray}
		\label{eq:rho_self_tunable}
		\rho = \frac{\vert y_{\mathrm{ref}(k)} \vert}{d_{\max}}.
	\end{eqnarray}

@@

	Note, the relations in Eq.~\eqref{eq:rho_self_tunable} and Eq.~\eqref{eq:rho_max} operate with the absolute value of the reference. It is not taken into account, whether the reference change was positive or negative with respect to the system steady-state value placed in the origin, i.e., whether it changed upwards or downwards. As many plants have asymmetric behavior, the positivity or negativity of the reference change could be considered in the controller tuning to improve the control performance.

@@

	\subsection{Tuning parameter based on the size of reference change}
	\label{sec:self_tunable_delta_ref}
	The approach of self-tunable explicit MPC in~\cite{self_tunable} suggested tuning based on the current reference value distance from the steady state. The aim is to compensate for the nonlinear behavior of a system when using a simple linear prediction model. This work provides also another useful way of the real-time evaluation of the tuning parameter $\rho$ based on the size of reference change. When different sizes of reference step changes are made and the behavior of the closed-loop system is varying, it can be beneficial to include the size of the reference step change in the tuning procedure.

@@

	\subsection{Self-tunable technique for systems with asymmetric behavior}
	\label{sec:self_tunable_rho_scaling}

	This paper provides a further extension of the self-tuning method proposed in~\cite{self_tunable}. The suggested technique of tuning is suitable, e.g., for systems with asymmetric behavior, but can be used in any application, where ``simple'' tuning in the whole range of tuning parameter $\rho$ is not sufficient.

@@

	%% TODO: DEF RULE/CONDITION GAMMA
	\begin{definition}[Decision function]
		\label{def:gamma}
		For a given interval of tuning parameter $\rho$, $0 \leq \rho \leq 1$, let $\rho_{\mathrm{s}}$, $0 < \rho_{\mathrm{s}} < 1$ be a boundary value splitting the interval into two parts. Let $\gamma: \mathbb{R} \rightarrow \mathbb{R}$ be an arbitrary function such that $0 \leq \gamma \leq 1$ holds. Then the decision function $\gamma$ is constructed to assign its value either $\gamma \leq \rho_{\mathrm{s}}$ or $\gamma > \rho_{\mathrm{s}}$.
	\end{definition}
	\textcolor{red}{TODO: replace in the line above: $\gamma \ge \rho_{\mathrm{s}}$?}\\
	Various decision functions $\gamma$ can be considered. In this work, the decision functions according to Eq.~\eqref{eq:rho_max} and Eq.~\eqref{eq:rho_delta_ref} are considered, while Eq.~\eqref{eq:rho_delta_ref} was implemented in the experimental case study.

@@

	For a better insight into the proposed technique, the procedure of self-tuning is depicted in Fig.~\ref{fig:tuning}.
